---
title: "Exercise 2"
subtitle: "Exercises for Exam 2"
categories: "Exercise"
format:
  html:
    mathjax: true
    html-math-method: mathjax
editor: source
execute: 
  include: true
---

(@) An independent random sample is selected from an approximately normal population with an unknown standard  deviation. Find the p-value for the given sample size and test statistic. Also determine if the null hypothesis of a two-tailed test would be rejected at $\alpha = 0.05$.
    (a) $n = 11$, $t_{test} = 1.91$  
    (b) $n = 17$, $t_{test} = -3.45$




```{r, echo=FALSE, eval=TRUE}
## 1.
cat("a) p-value=", 2 * pt(q = 1.91, df = 11 - 1, lower.tail = FALSE))
## 2.
cat("b) p-value=",  2 * pt(q = -3.45, df = 17 - 1, lower.tail = TRUE))
```
<br>




(@) Georgianna claims that in a small city renowned for its music school, the average child takes less than 5 years of piano lessons. We have a random sample of 20 children from the city, with a mean of 4.6 years of piano lessons and a standard deviation of 2.2 years.
    (a) Evaluate Georgianna's claim (or that the opposite might be true) using a hypothesis test.
    (b) Construct a 95\% confidence interval for the number of years students in this city take piano lessons, and interpret it in context of the data.
    (c) Do your results from the hypothesis test and the confidence interval agree? Explain your reasoning.


```{r, echo=TRUE, eval=TRUE}
## a) One sample t test with alpha 0.05
## H0: mu >= 5; H1: mu < 5
(t_test <- (4.6 - 5) / (2.2/sqrt(20)))
(t_cri <- qt(p = 0.05, df = 20 - 1)) ## Do not reject H0


## b) 
4.6 + c(-1, 1) * qt(p = 0.975, df = 20 - 1) * (2.2 / sqrt(20))


## c)
# Not agree. The test is one-sided, but the CI is two-sided.
```

```{r, include=FALSE}
# OI 7.11 Play the piano
```

<br>



(@) In each of the following scenarios, determine if the data are paired.
    (a) Compare pre- (beginning of semester) and post-test (end of semester)  scores of students.
    (b) Assess gender-related salary gap by comparing salaries of randomly  sampled men and women.
    (c) Compare artery thicknesses at the beginning of a study and after 2 years  of taking Vitamin E for the same group of patients.
    (d) Assess effectiveness of a diet regimen by comparing the before and after  weights of subjects.

```{r, echo=FALSE, eval=TRUE}
cat("a) paired. ")
cat("b) independent. ")
cat("c) paired. ")
cat("d) paired. ")
```


```{r, include=FALSE}
## OI 7.17
# (a) Since it's the same students at the beginning and the end of the semester, there is a pairing between the datasets, for a given student their beginning and end of semester grades are dependent.
# (b) Since the subjects were sampled randomly, each observation in the men's group does not have a special correspondence with exactly one observation in the other (women's) group.
# (c) Since it's the same subjects at the beginning and the end of the study, there is a pairing between the datasets, for a subject student their beginning and end of semester artery thickness are dependent.
# (d) Since it's the same subjects at the beginning and the end of the study, there is a pairing between the datasets, for a subject student their beginning and end of semester weights are dependent.
```

<br>


(@) Dr. Yu decided to run two slight variations of the same exam. Prior to passing out the exams, he shuffled the exams together to ensure each student received a random version. Summary statistics for how students performed on these two exams are shown below. Anticipating complaints from students who took Version B, he would like to evaluate whether the difference observed in the groups is so large that it provides convincing evidence that Version B was more difficult (on average) than Version A. Test the claim with $\alpha = 0.01$.
<!-- \begin{figure}[hht] -->
<!-- \centering -->
<!-- \begin{tabular}{l rrrrr} -->
<!-- \hline -->
<!-- Version\hspace{2mm}	& $n$	& $\bar{x}$	& $s$	& min	& max  \\ -->
<!-- \hline -->
<!-- A		& 30		& 79.4		& 14 	& 45		& 100 \\ -->
<!-- B		& 27		& 74.1		& 20		& 32		& 100 \\ -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \end{figure} -->

```{r}
#| echo: false
library(gt)

tbl <- data.frame(
Version = c("A", "B"),
n       = c(30, 27),
xbar    = c(79.4, 74.1),
s       = c(14, 20),
min     = c(45, 32),
max     = c(100, 100)
)

tbl |>
gt() |>
cols_label(
n    = html("\\(n\\)"),
xbar = html("\\(\\bar{x}\\)"),
s    = html("\\(s\\)"),
min  = "min",
max  = "max"
)

```


```{r, echo=c(-1)}
## OI 7.3.3 Case study: two versions of a course exam
n1 = 30; x1_bar = 79.4; s1 = 14
n2 = 27; x2_bar = 74.1; s2 = 20
A <- s1^2 / n1; B <- s2^2 / n2
df <- (A + B)^2 / (A^2/(n1-1) + B^2/(n2-1))
(df <- floor(df))
## t_test
(t_test <- (x1_bar - x2_bar) / sqrt(s1^2/n1 + s2^2/n2))
## t_cv
qt(p = 0.01, df = df, lower.tail = FALSE)
## p_value
pt(q = t_test, df = df, lower.tail = FALSE)
# the data do not convincingly show that one exam
# version is more difficult than the other
```

<br>  
<br>  


(@) Undergraduate students taking an introductory statistics course at Marquette University conducted a survey about GPA and major. The ANOVA output is provided.
    (a) Write the hypotheses for testing for a difference between average GPA across majors.
    (b) What is the conclusion of the hypothesis test?
    (c) How many students answered these questions on the survey, i.e. what is the sample size?
```{r}
#| echo: false
library(gt)

anova_tbl <- data.frame(
Source   = c("major", "Residuals"),
Df       = c(2, 195),
SumSq    = c(0.03, 15.77),
MeanSq   = c(0.015, 0.081),
Fvalue   = c(0.185, ""),
PrF      = c(0.8313, "")
)

anova_tbl |>
gt() |>
cols_label(
Source = "Source",
Df     = "Df",
SumSq  = "Sum Sq",
MeanSq = "Mean Sq",
Fvalue = "F value",
PrF    = html("Pr(>F)")
)
```

```{r, echo=c(-1)}
## OI 7.41
## (a) 
# H0: mu1 = mu2 = mu3; H1: not all mus are equal

## (b)
# p-value > 0.05. The data do not provide convincing evidence of a
# difference between the average GPAs across three groups of majors.

## (c)
# The total degrees of freedom is 195 + 2 = 197, so the sample size is 197 + 1 = 198.
```


<br>  
<br>  


(@) Find the test statistic, critical value(s) of $\chi^2$, and p-value, then determine whether there is sufficient evidence to support the given alternative hypothesis.
    (a) $H_1: \sigma \ne 15$, $\alpha = 0.05$, $n = 20$, $s = 10$.
    (b) $H_1: \sigma > 12$, $\alpha = 0.01$, $n = 5$, $s = 18$.

```{r, echo=c(-1)}
## Bio 7.6
## (a)
## test statistic
(20 - 1)*10^2/(15^2)
## critical values
qchisq(p = 0.05/2, df = 19, lower.tail = TRUE)
qchisq(p = 0.05/2, df = 19, lower.tail = FALSE)
## p-value
2*pchisq((20 - 1)*10^2/(15^2), df = 19)

## (b)
## test statistic
(5 - 1)*18^2/(12^2)
## critical value
qchisq(p = 0.01, df = 4, lower.tail = FALSE)
## p-value
pchisq((5 - 1)*18^2/(12^2), df = 4, lower.tail = FALSE)
```

(@) A patient named Diana was diagnosed with Fibromyalgia, a long-term
syndrome of body pain, and was prescribed anti-depressants. Being the skeptic that she is, Diana didn't initially believe that anti-depressants would help her symptoms. However after a couple months of being on the medication she decides that the anti-depressants are working, because she feels like her symptoms are in fact getting better.
    (a) Write the hypotheses in words for Diana's skeptical position when she started taking the anti-depressants.
    (b) What is a Type 1 Error in this context?
    (c) What is a Type 2 Error in this context?
```{r, include=TRUE, echo=-1}
## OI 5.25
## (a)
## H0: Anti-depressants do not affect the symptoms of Fibromyalgia. 
## HA: Anti-depressants do affect the symptoms of Fibromyalgia (either helping or harming).
## (b)
## Concluding that anti-depressants either help or worsen Fibromyalgia symptoms 
## when they actually do neither.
## (c)
## Concluding that anti-depressants do not affect Fibromyalgia symptoms when they actually do.
```


(@) **True or false**.
    (a) The chi-square distribution, just like the normal distribution, has two parameters, mean and standard deviation.
    (b) Consider two sets of data that are paired with each other. Each observation in one data set has a natural correspondence with exactly one observation from the other data set.
    (c) As the degrees of freedom increases, the $t$-distribution approaches normality.
    (d) A correlation coefficient of -0.90 indicates a stronger linear relationship than a correlation of 0.5.

```{r, include=TRUE, echo=TRUE}
## F
## T
## T
## T
```

(@) Consider a regression predicting weight (kg) from height (cm) for a sample of adult males. What are the units of the correlation coefficient, the intercept, and the slope?

```{r, include=TRUE, echo=-1}
## OI 8.17
# Correlation: no units. Intercept: kg. Slope: kg/cm.
```


(@) Is the gestational age (time between conception and birth) of a low birth-weight baby useful in predicting head circumference at birth? Twenty-five low birth-weight babies were studied at a Harvard teaching hospital; the investigators calculated the regression of head circumference (measured in centimeters) against gestational age (measured in weeks). The estimated regression line is $$\widehat{\text{head circumference}} = 3.91 + 0.78 \times \text{gestational age}$$
    (a) What is the predicted head circumference for a baby whose gestational age is 28 weeks?

```{r, include=TRUE, echo=-1}
## OI 8.36
# 3.91 + 0.78 * 28 = 25.75
```
