---
title: "Homework 2"
subtitle: "Basic R, Data Description and Graphics"
categories: "Homework"
editor: source
execute:
  echo: false
  eval: false
---

::: callout-important
[Due **Friday, Feb 6, 11:59 PM**]{.red}
:::

- Homework 2 covers Week 2 and 3 course materials.

- Please submit your work in *one* **PDF** file to **D2L > Assessments > Dropbox**. Multiple files or a file that is not in pdf format is not accepted. 

- **Attach your R code** to show how you obtain your answer. **\color{black}{If not, you receive no score.}**

- **Hand drawn tables and graphs receive no credits.**

<!-- - Your homework can be handwritten or typewritten. If it is handwritten, make sure you scan your work well, and put everything in one single pdf. -->

<!-- - If you type your answers, make sure there are no typos. I grade your work based on *what you show, not what you want to show*. If you choose to handwrite your answers, write them neatly. If I can’t read your handwriting, your answer is judged as wrong. -->

- Questions started with **(MSSC)** are *required* for MSSC 5720 students, and optional for MATH 4720 students.


# Homework Questions


```{r}
#| echo: true
#| eval: true
# ==============================================================================
## Vector
# ==============================================================================
poker_vec <- c(140, -50, 20, -120, 240)
roulette_vec <- c(-24, -50, 100, -350, 10)
days_vec <- c("Mon", "Tue", "Wed", "Thu", "Fri")
names(poker_vec) <- days_vec
names(roulette_vec) <- days_vec
```

(@) **Vector.** The code above shows a Marquette student poker and roulette winnings from Monday to Friday. Copy and Paste them into your R and complete the problem 1.
    (a) Assign to the variable `total_daily` how much you won or lost on each day in total (poker and roulette combined).
    (b) Calculate the winnings overall `total_week`. Print it out.

```{r}
#| echo: true
#| eval: true
## (a)
total_daily <- poker_vec + roulette_vec
total_daily

## (b)
total_week <- sum(total_daily)
total_week
```

<br>

```{r}
#| echo: true
#| eval: true
# ==============================================================================
## Factor
# ==============================================================================
# Create speed_vector
speed_vec <- c("medium", "slow", "slow", "medium", "fast")
```


(@) **Factor.**
    (a) `speed_vec` above should be converted to an **ordinal** factor since its categories have a natural ordering. Create an ordered factor vector `speed_fac` by completing the code below. Set the argument `ordered` to `TRUE`, and set the argument `levels` to `c("slow", "medium", "fast")`. Print `speed_fac`.


```{r}
#| echo: true
#| eval: true
## (a)
(speed_fac <- factor(speed_vec, ordered = TRUE, 
                    levels = c("slow", "medium", "fast")))
```

<br>

```{r}
#| echo: true
#| eval: true
# ==============================================================================
## Data frame
# ==============================================================================
name <- c("Mercury", "Venus", "Earth", "Mars", "Jupiter", "Saturn", 
          "Uranus", "Neptune")
type <- c("Terrestrial planet", "Terrestrial planet", "Terrestrial planet", 
          "Terrestrial planet", "Gas giant", "Gas giant", 
          "Gas giant", "Gas giant")
diameter <- c(0.382, 0.949, 1, 0.532, 11.209, 9.449, 4.007, 3.883)
rotation <- c(58.64, -243.02, 1, 1.03, 0.41, 0.43, -0.72, 0.67)
rings <- c(FALSE, FALSE, FALSE, FALSE, TRUE, TRUE, TRUE, TRUE)
```

(@) **Data Frame.** You want to construct a data frame that describes the main characteristics of eight planets in our solar system. You feel confident enough to create the necessary vectors: `name`, `type`, `diameter`, `rotation` and `rings` that have already been coded up as above. The first element in each of these vectors correspond to the first observation.
    (a) Use the function `data.frame()` to construct a data frame. Pass the vectors `name`, `type`, `diameter`, `rotation` and `rings` as arguments to `data.frame()`, in this order. Call the resulting data frame `planets_df`.
    (b) Use `str()` to investigate the structure of the new `planets_df` variable. Which are categorical (qualitative) variables and which are numerical (quantitative) variables? For those that are categorical, are they nominal or ordinal? For those numerical variables, are they interval or ratio level? discrete or continuous?
    (c) From `planets_df`, select the diameter of Mercury: this is the value at the first row and the third column. Simply print out the result.
    (d) From `planets_df`, select all data on Mars (the fourth row). Simply print out the result.
    (e) Select and print out the first 5 values in the `diameter` column of `planets_df`.
    (f) Use `$` to select the `rings` variable from `planets_df`.
    <!-- (g) Use (f) to select all columns for planets that have rings. -->


```{r}
#| eval: true
#| echo: true
## (a)
planets_df <- data.frame(name, type, diameter, rotation, rings)
planets_df

## (b)
str(planets_df)
# name, type, rings: categorical, nominal
# diameter, rotation: numerical, ratio, continuous


## (c)
planets_df[1, 3]

## (d)
planets_df[4, ]


## (e)
planets_df[1:5, "diameter"]

## (f)
planets_df$rings
```

```{r}
## (g)
planets_df[planets_df$rings, ]
```

<br>   

(@) **Data Description and Graphics.** We use the data set `mtcars` to do data summary and graphics. Type `?mtcars` for the description of the data set.
    (a) Use the function `pie()` to create a pie chart for the number of cylinders (`cyl`). Show the plot. What the number of cylinders has the most frequencies in the data?
    (b) Use the function `barplot()` to create a bar chart for the number of gears (`gear`). Show the plot. What the number of gears has the most frequencies in the data?
    (c) Use the function `hist()` to generate a histogram of the gross horsepower (`hp`). Show the plot. Is it right or left-skewed?
    (d) Use the function `boxplot()` to generate a boxplot of car weight (`wt`). Show the plot. Are there any outliers?
    (e) Use the function `plot()` to create a scatter plot of displacement (`disp`) vs. miles per gallon (`mpg`). Show the plot. As the displacement increases, the miles per gallon tends to increase or decrease?
    <!-- (f) Compute the mean, median and standard deviation of the miles per gallon (`mpg`). -->
    <!-- (a) Use the function `pie()` to create a pie chart for the number of cylinders (`cyl`). What the number of cylinders has the most frequencies in the data? -->
    <!-- (b) Use the function `barplot()` to create a bar chart for the number of gears (`gear`). What the number of gears has the most frequencies in the data? -->
    <!-- (c) Use the function `hist()` to generate a histogram of the gross horsepower (`hp`). Is it right or left-skewed? -->
    <!-- (d) Use the function `boxplot()` to generate a boxplot of car weight (`wt`). Are there any outliers? -->
    <!-- (e) Use the function `plot()` to create a scatter plot of displacement (`disp`) vs. miles per gallon (`mpg`). As the displacement increases, the miles per gallon tends to increase or decrease? -->
    <!-- (f) Compute the mean, median and standard deviation of the miles per gallon (`mpg`). -->


```{r}
#| echo: true
#| eval: true
## (a)
## most cars have 8 cylinders
pie(table(mtcars$cyl))
## (b)
## most cars have 3 gears
barplot(table(mtcars$gear))
## (c)
## right-skewed
hist(mtcars$hp, breaks = 10)
## (d) 
## there are two outliers
boxplot(mtcars$wt)
## (e)
## As the displacement increases, the miles per gallon tends to decrease.
plot(x = mtcars$disp, y = mtcars$mpg)
```

To save your figures, in RStudio you go to tab **Plots > Export > Save as Image > choose Image format (PNG is good), choose where the image is saved (Directory), type the File name, decide Width and Height > Click Save**. To download an image file to your local computer, select the file, go to **More > Export > Download**. You could take screenshots of plots or code to show your work too.

```{r}
## (f)
mean(mtcars$mpg)
median(mtcars$mpg)
sd(mtcars$mpg)
```

<br>

(@) **(MSSC) R List Data Structure.** A **list** in R allows you to gather a variety of objects under one name (that is, the name of the list) in an ordered way. These objects can be matrices, vectors, data frames, even other lists. Alos, it is not required that these objects are related to each other in any way.
    (a) Use command  `list()` to create a list named `employee` of three elements, `name`, `salary`, and `union` with value `"Joe"`, `55000` and `TRUE` respectively. Print `employee`.
    (b) Construct a list, named `my_list` that contains 3 list components: `days_vec`, `planets_df`, and `employee`. Print the structure of `my_list`.  
    (c) What is the difference between `my_list[[1]]` and `my_list[1]`? What is the data type they return?
    (d) Obtain the salary in `employee` from `my_list` using `$`.


```{r}
#| echo: true
#| eval: true
## (a)
employee <- list(name = "Joe", salary = 55000, union = TRUE)
employee
## (b)
my_list <- list(days_vec, planets_df, employee)
my_list
## (c)
## my_list[[1]] returns a vector of type character
my_list[[1]]
typeof(my_list[[1]])
## my_list[1] keeps the list structure with the first element of my_list
my_list[1]
typeof(my_list[1])
## (d)
my_list[[3]]$salary
```

<br>

(@) **AI Usage Declaration.** Using GenAI is permitted for this course. If you choose to use GenAI to assist with your homework, you must include a brief statement documenting your use. Please provide the following information:
    (a) **Why/How I Used AI** Why do you need to use GenAI? Which tool did you use? Describe your prompts or questions. What and how did you ask the AI to help you?
    (b) **Generated Output** Include a screenshot or excerpt (copy and paste) of the AI’s response.
    (c) **How I Used the Output** Did you revise it? Did you use it directly, or compare it with your answers? What decisions did you make based on the output?

