---
title: "Activity: Descriptive statistics"
execute:
  echo: true
---

## **Objective**

1. Create and interpret distribution graphics.
2. Compute mean, median, IQR, and range in base R, including handling missing values.
3. Identify outliers and decide which summaries best describe typical and variability.


## Instructions

Form a group of **four** or **five**.


### Load `airquality` data

```{r}
data(airquality)
head(airquality)
str(airquality)
```

::: {.callout-note}
## Discussion

- What does the function `head()` do?
- What does `NA` mean?
- Discuss the data type of the variables in the `airquality` data. 

:::

### Create the histogram and the boxplot of the variable `Ozone`

```{r}
#| eval: false
____(_________________, breaks = 15, main = "Histogram of Ozone")
```

```{r}
#| eval: false
_______(_________________, main = "Boxplot of Ozone")
```

```{r}
#| eval: false
#| echo: false
hist(airquality$Ozone, breaks = 15, 
     main = "Histogram of Ozone", xlab = "Ozone (ppb)")
```

::: {.callout-note}
## Discussion

- Is the distribution symmetric, left-skewed or right-skewed?
- Is the distribution unimodal or bimodal?
- Are there outliers?

:::

### Boxplot of `ozone` by `month`

```{r}
#| eval: false
boxplot(_____ ~ _____, data = airquality)
```

::: {.callout-note}
## Discussion

- Which month has the highest typical ozone?
- Which month looks most variable?

:::


## Obtain the mean, median, IQR, minimum, and maximum of `Ozone`

::: {.callout-note}
## Discussion

- Which is larger, the mean or the median?
- What does that suggest about skewness and outliers?

:::


## Conclusion

Fill in the blanks using your results.

> 1. Typical ozone: Month __ has the highest typical ozone because its boxplot median is highest, and the median ozone is __ (ppb).

> 2. Variability: Month __ is most variable because its boxplot is widest, and its IQR is __ (ppb).

> 3. Summary choice: Because the plots show outliers and skewness, we trust __ more than __ for describing a typical day and variability.


{{< countdown minutes=10 font-size="5rem" >}}
